---
title: "Longhurst maps"
author: "Andrew Irwin"
date: '2022-04-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(sp)
library(ggmap)
library(rnaturalearth)
library(rnaturalearthdata)
# discussion of slow map plotting in Mac versions
# https://github.com/tidyverse/ggplot2/issues/2655
```

Read shapefiles

```{r}
longhurst <- st_read("shapefile/Longhurst_world_v4_2010.shp")
longhurst$ProvCode  = as.character(longhurst$ProvCode)
longhurst$ProvCode[longhurst$ProvCode == "PEQD"] = "PEOD"  # typo
```

Get map data.

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
# world <- ne_countries(scale = "medium", returnclass = "sf", type="map_units")
longhurst_sf <- st_sf(longhurst)
# longhurst_4326 <- st_transform(longhurst, 4326)
```

Pick a few regions.

```{r}
p0 <- ggplot(data = world) + geom_sf(fill="#A0A0A0", color="#808080") +
  geom_sf(data = longhurst[52,], size = 1, fill = "red", color = NA) + 
  geom_sf(data = longhurst[40:50,], size = 1, fill = "black", color = NA) + 
  geom_sf(data = longhurst[51,], size = 1, fill = "blue", color = NA) + 
  ggtitle("Longhurst") + 
  coord_sf() +
  theme_bw() + 
  theme(panel.border =  element_blank()) + 
  # coord_sf(crs = "+proj=laea +lat_0=-90 +lon_0=0") # S pole
  # coord_sf(crs = "+proj=ortho +lat_0=90") # N pole
  # coord_sf(crs = "+proj=moll +lon_0=0") # Mollwiede, globe
  coord_sf(crs = "+proj=eck4")
  # coord_sf(crs = 4326)
ggsave('map0.pdf', p0)
ggsave('map0.png', p0)
```

Zoom in on a region.

```{r}
p2 <- ggplot(data = world) + geom_sf(fill="#A0A0A0", color="#808080") +
  # geom_sf(data = longhurst2 %>% mutate(geometry = st_transform(geometry, "+proj=eck4 +lat_0=+0 +lon_0=0")),
  geom_sf(data = longhurst2,
          size = 0.5, color = NA, aes(fill=ProvCode)) + 
  scale_fill_brewer(palette = "Set3") + 
  # coord_sf(crs = "+proj=eck4") +  # projecting and cropping a map is hard...
  coord_sf(crs = 4326) +
  xlim(-80,10) + ylim(25, 75) +
  # coord_sf(crs = "+proj=eck4 +lat_0=+0 +lon_0=0") # slow and fails.
  theme_void()
ggsave("map2.pdf", p2)

```

Project and zoom using st_crop. 
See https://www.r-bloggers.com/2019/04/zooming-in-on-maps-with-sf-and-ggplot2/

```{r}
# w2 <- world %>% st_crop(ymin = 25, ymax = 75, xmin = -80, xmax = 10)  # spherical geometry errors
crop_region <- st_sfc(st_point(c(-80, 25)), st_point(c(10, 75)), crs = 4326)
target_crs = "+proj=eck4 +lat_0=+00 +lon_0=0"  
# eck4, moll have trouble with lines breaking at edges depending on lon_0 
crop_region_t <- st_transform(crop_region, crs = target_crs)
crop_region_c <- st_coordinates(crop_region_t)
p2 <- ggplot(data = world) + 
  geom_sf(fill="#A0A0A0", color=NA) + # "#808080") +
  geom_sf(data = longhurst2,
          size = 0.5, color = NA, aes(fill=ProvCode)) + 
  scale_fill_brewer(palette = "Set3") + 
  coord_sf(crs = target_crs, xlim = crop_region_c[,"X"],
                             ylim = crop_region_c[,"Y"]) +  
  theme_void()
ggsave("map2.pdf", p2)

```

Add points to a plot. Not working.

```{r}
db <- tibble(latitude = runif(20, 20, 75),
             longitude = runif(20, -80, 20))
```


```{r}
longhurst_t <- st_transform(longhurst, 4326)
# points1_t <- st_transform(as.matrix(db[1:10, c("longitude", "latitude")]), 2163)  # doesn't work
tidyLatLong <- function(db) {
  points1 <- db[ , c("longitude", "latitude")]
  names(points1) <- c("x", "y")
  points1_sf <- do.call("st_sfc",
                        c(lapply(1:nrow(points1), function(i) {st_point(as.numeric(points1[i, ]))}), list("crs" = 4326))) 
  # points1_t <- st_transform(points1_sf, 4326)
  points1_sf
}
points1_t <- tidyLatLong(db)
db %>% cbind(points1_t) %>%
  mutate(longhurst = apply(
    st_intersects(longhurst_t, geometry, sparse=FALSE),
    2,
    function(col) paste0(longhurst_t$ProvCode[which(col)], collapse=",")
  )) -> db
```


```{r}
ggplot(data = world) + geom_sf(fill="#E0E0E0", color="#E0E0E0") +
  geom_sf(data = db %>% distinct(latitude, longitude, .keep_all = TRUE), # %>% sample_n(100),
          aes(geometry = geometry_eck4, color=longhurst)) +
  geom_sf(data = longhurst, fill=NA) +
  coord_sf(crs = "+proj=eck4 +lat_0=+0 +lon_0=0") +
    theme_bw() -> p
ggsave("map_points.pdf", p)
```





## Classify points

Move to separate file.
Improve code.


How to determine which region a particular lat, long belongs to?

```{r}
points1 <- st_multipoint(as.matrix(db[1:10, c("longitude", "latitude")]), "XY")
longhurst_sf <- st_sf(longhurst)
longhurst_sf %>% select(ProvCode, ProvDescr)
st_within(longhurst, points1)
```

WGS84 (world geodetic system 1984) has CRS code 4326
US equal area atlas has CRS code 2163
Pseudo mercator has CRS code 3857 and used by google maps, etc.

```{r}
longhurst_t <- st_transform(longhurst, 4326)
# points1_t <- st_transform(as.matrix(db[1:10, c("longitude", "latitude")]), 2163)  # doesn't work
tidyLatLong <- function(db) {
  points1 <- db[ , c("longitude", "latitude")]
  names(points1) <- c("x", "y")
  points1_sf <- do.call("st_sfc",
                        c(lapply(1:nrow(points1), function(i) {st_point(as.numeric(points1[i, ]))}), list("crs" = 4326))) 
  # points1_t <- st_transform(points1_sf, 4326)
  points1_sf
}
points1_t <- tidyLatLong(db)
db %>% cbind(points1_t) %>%
  mutate(longhurst = apply(
    st_intersects(longhurst_t, geometry, sparse=FALSE),
    2,
    function(col) paste0(longhurst_t$ProvCode[which(col)], collapse=",")
  )) -> db
```




